### 2018-03-13 submission/manuscript.tex
all pat's changes except fig adjustment and RF section

\textbf{\emph{Performance of models based on OTU relative abundances are
not significantly better than those based on taxa with significant
ORs.}} The previous models were based on relative abundance data where
sequences were classified to coarse taxonomic assignments
(i.e.~typically genus or family level). To determine whether model
performance improved with finer scale classification, we assigned
sequences to operational taxonomic units (OTUs) where the similarity
among sequences within an OTU was more than 97\%. We again found that
classification models built using all of the sequence data for a
community did a poor job of differentiating between subjects with normal
colons and those with adenomas (median AUC: 0.53 (0.37- 0.56)). However,
they did a good job of differentiating between subjects with normal
colons and those with carcinomas (median AUC: 0.71 (0.50- 0.904)). The
OTU-based models performed similarly to those constructed using the taxa
with significant ORs (one-tailed paired T-test; P-value = 0.966) and
those using all taxa (one-tailed paired T-test; P-value = 0.146)
{[}Figure 4{]}. Among the OTUs that had the highest mean decrease in
accuracy for the OTU-based models, we found that OTUs that affiliated
with all of the 8 taxa that had a significant OR were within the top
10\% for at least one study. This result was surprising as it indicated
that a finer scale classification of the sequences and thus a larger
number of features to select from, did not yield improved classification
of the subjects.

\textbf{\emph{Generalizability of taxon-based models trained on one
dataset to the other datasets.}} Considering the good performance of the
Random Forest models using taxa with a significant OR and using all of
the taxa, we next asked how well the models would perform when given
data from a different subject cohort. For instance, if a model was
trained using data from the Ahn study, we wanted to know how well it
would perform using the data from the Baxter study. We found the models
trained using the taxa with a significant OR all had a higher median AUC
than the models trained using all of the taxa when tested on the other
datasets {[}Figure 6 \& S5{]}. As might be expected, the difference
between the performance of the modelling approaches appeared to vary
with the size of the training cohort (R\textsuperscript{2} = 0.66)
{[}Figure 6{]}. These data suggest that given a sufficient number of
subjects with normal colons and carcinomas, Random Forest models trained
using a small number of taxa can accurately classify individuals from a
different cohort.

@@

We performed a meta-analysis to identify and validate microbiome-base
biomarkers that could be used to classify individuals as having normal
colons or colonic tumors using fecal or tissue samples. To our surprise,
Random Forest classification models constructed to differentiate
individuals with normal colons from those with carcinomas using a subset
of the community performed well relative to models constructed using the
full communities. When we applied the models trained on each dataset to
the other datasets in our study, we found that the models trained using
the subset of the communities performed better than those using the full
communities. These models were trained using data in which sequences
were assigned to bacterial taxa using a classifier that typically
assigned sequences to the family or genus level. When we attempted to
improve the specificity of the classification by using an OTU-based
approach the resulting models performed as well as those constructed
using coarse taxonomic assignments. These results are significant
because they strengthen the growing literature indicating a role of the
microbiome in tumorigenesis (9) and as a potential tool as a
non-invasive diagnostic and for assessing risk of disease and recurrence
(12, 30).

These results suggested that fine scale classification of sequences into
OTUs does not improve our classification models. This has been suggested
in previous literature where shotgun metagenomic data did not perform
better than 16S rRNA gene sequencing data in classifying individuals
with normal colons and those with carcinomas (31). We hypothesize that
fine scale classification may not result in better classification
because distribution of microbiota between individuals is patchy. In
contrast, models using coarser taxonomic assignments will pool the fine
scale diversity, resulting in less patchiness and better classification.
Furthermore, the ability of models trained using a subset of the
community to outperform those using the full community when testing the
models on the other datasets may also be a product of the patchiness of
the human-associated microbiota. The models based on the 8 taxa that had
significant ORs used taxa that were found in every study and tended to
have higher relative abundances. Similar to the OTU-based models, those
models based on the full community taxonomy assignments were still
sensitive to the patchy distribution of taxa. Regardless, it is
encouraging that a collection of 8 taxa could reliably classify
individuals as having carcinomas considering the differences in cohorts,
DNA extraction procedures, regions of the 16S rRNA gene, and sequencing
methods.

When used separately to classify individuals with carcinomas, the taxa
with significant ORs could not reliably classify individuals {[}Figure
3{]}. This result further supports the hypothesis that
carcinoma-associated microbiota have a patchy distribution. Two
individuals may have had the same classification, based on the relative
abundance of different populations within this group of 8 taxa. Although
these results only reflect associations with disease, it is tempting to
hypothesize that the patchiness represents distinct mechanisms of
exacerbating tumorigenesis or that multiple taxa have the same mechanism
of exacerbating tumorigenesis. For example, strains of \emph{Escherichia
coli} and \emph{Fusobacterium nucleatum} have been shown to worsen
inflammation in mouse models of tumorigenesis (5, 6, 21). In contrast to
the patchiness of the taxa that were positively associated with
carcinomas, potentially beneficial taxa had a more consistent
association {[}Figure 6{]}. This result was particularly interesting
because members of these taxa (i.e. \emph{Ruminococcus} and
\emph{Clostridium XI} in stool and \emph{Dorea} and \emph{Blautia} in
tissue) are thought to be beneficial due to their involvement in
production of anti-inflammatory short chain fatty acids (32--34).

All of the adenoma classification models performed poorly and is not
inconsistent with previous studies (27, 30). However, the classification
results are at odds with results of the multitarget microbiota test
(MMT) from Baxter, \emph{et al.} (12) who observed an AUC of 0.755 when
applied to individuals with adenomas. There are two major differences
between the models generated in this meta-analysis and that analysis.
The MMT attempted to classify individuals as having a normal colon or
having colonic lesions (i.e.~adenomas or carcinomas) and not adenomas
alone. Further, the MMT incorporated fecal immunoglobulin test (FIT)
data while our models only used 16S rRNA gene sequencing data. Because
FIT data were not available for the other studies in our meta-analysis,
it was not possible to validate the MMT approach. The ability to
differentiate between individuals with and without adenomas is an
important problem since early detection of tumors is critical. However,
it is possible that we might have been able to detect differences in the
bacterial community if individuals with non-advanced and advanced
adenomas were separated. This is a clinically relevant distinction since
advanced adenomas are at highest risk of progressing to a carcinoma. The
initial changes of the microbiota during tumorigenesis could be focal to
where the initial adenoma develops and would not be easily assessed
using fecal samples from an individual with non-advanced adenomas.
Unfortunately, distinguishing between individuals with advanced and
non-advanced adenomas was not possible in our meta-analysis since the
studies did not provide the clinical data needed to make that
distinction.

Stool samples represent a non-invasive approach to assessing the
structure of the gut microbiota and are potentially useful for
diagnosing individuals as having colonic tumors. However, they do not
reflect the structure of the mucosal microbiota (35). Regardless, the
taxa that were the most important in the stool-based models overlapped
with those from the models trained using the data from unmatched and
matched colon tissue samples {[}Figure S3{]}. Mucosal biopsies are
preferred for focused mechanistic studies and have offered researchers
the opportunity to sample healthy and diseased tissue from the same
individuals (i.e.~matched) using each individual as their own control or
in a cross sectional design (i.e.~unmatched). Because obtaining these
samples is invasive, carries risks to the individual, and is expensive,
studies investigating the structure of the mucosal microbiota generally
have a limited number of participants. Thus, it was not surprising that
tissue-based studies did not provide clearer associations between the
mucosal microbiota and the presence of tumors. Interestingly,
Fusobacterium, which has received increased recent attention for its
potential role in tumorigenesis (6) was not consistently identified
across the studies in our meta-analysis. This could be due to the
relatively small number of individuals in the limited number of studies.
The classification models trained using the tissue-based data performed
well when tested with the training data (Figure S4), but performed
poorly when tested on the other tissue-associated datasets (Figure S5).
Disturbingly, taxa that are commonly associated with reagent
contamination (e.g. \emph{Novosphingobium}, \emph{Acidobacteria Gp2},
\emph{Sphingomonas}, etc.) were detected within the tissue datasets.
Such contamination is common in studies where there is relatively low
bacterial biomass (36). The lack of replication among the tissue-based
biomarkers may be a product of the relatively small number of studies
and individuals per study and possible reagent contamination.

Among our stool samples, we failed to identify several notable
populations that are commonly associated with carcinomas including an
enterotoxigenic strain of \emph{Bacteroides fragilis} (ETBF) and
\emph{Streptococcus gallolyticus} subsp. \emph{gallolyticus} (22, 24).
ETBF have been found in tumors in the proximal colon where they tend to
form biofilms (20, 37). Considering DNA from bacteria that are more
prevalent in the proximal colon may be degraded by the time it leaves
the body, it is not surprising that we failed to identify a significant
OR for Bacteroides with carcinomas. In addition, since our approach
could only classify sequences to the genus level and there are likely
multiple Bacteroides populations in the colon, it is possible that
sequences from ETBF and non-oncogenic Bacteroides were pooled. This
would then reduce the OR between Bacterioides and whether an individual
had carcinomas. It is also necessary to distinguish between populations
that are biomarkers for a disease and those that are known to cause
disease. The former may also have a causative role in the disease.
Although the latter have been shown to have a causative role, they may
appear at low relative abundance, be found in specific locations, or may
have a highly patchy distribution among affected individuals.

Meta-analyses are a useful tool in microbiome research because they can
demonstrate whether a result can be replicated and facilitate new
discoveries by pooling multiple independent investigations. There have
been several meta-analyses similar to this study that have sought
biomarkers for obesity (38--40), inflammatory bowel disease (39), and
colorectal cancer (28). Considering microbiome research is particularly
prone to hype and overgeneralization of results, these analyses are
critical. For example, previous meta-analyses have demonstrated that
there are no clear fecal biomarkers for obesity (39, 40). Such
meta-analyses are difficult to perform because the underlying 16S rRNA
gene sequence data are not publicly available, metadata are missing,
incomplete, or vague, sequence data are of poor quality or derived by
non-standard approaches, and the original studies were significantly
underpowered. Reluctance to publish negative results (i.e.~the ``file
drawer effect'') is also likely to skew our understanding of the
relationship between microbiota and disease. Better attention to these
specifics will increase the reproducibility and replicability of
microbiota studies and make it easier to perform these crucial
meta-analyses. Moving forward, meta-analyses will be important tools to
help aggregate and find commonalities across studies when investigating
the microbiota in the context of a specific disease (28, 38--40).

@@

\textbf{\emph{Datasets.}} The studies used for this meta-analysis were
identified through the review articles written by Keku, et al. (41) and
Vogtmann, et al. (42). Additional studies, not mentioned in those
reviews were obtained based on the authors' knowledge of the literature.
Studies were included that used tissue or feces as their sample source
for 454 or Illumina 16S rRNA gene sequencing. Some studies (N = 12) were
excluded because they did not have publicly available sequences, did not
use 454 or Illumina machines, or did not have metadata in which the
authors were able to share. We were able to obtain sequence data and
metadata from the following studies: Ahn, \emph{et al.} (11), Baxter,
\emph{et al.} (12), Brim, \emph{et al.} (29), Burns, \emph{et al.} (15),
Chen, \emph{et al.} (13), Dejea, \emph{et al.} (20), Flemer, \emph{et
al.} (17), Geng, \emph{et al.} (19), Hale, \emph{et al.} (27), Kostic,
\emph{et al.} (43), Lu, \emph{et al.} (26), Sanapareddy, \emph{et al.}
(25), Wang, \emph{et al.} (14), Weir, \emph{et al.} (23), and Zeller,
\emph{et al.} (16). The Zackular (44) study was excluded because their
90 individuals were contained within the larger Baxter study (12). The
Kostic study was excluded because after we processed the sequences, all
of the case samples had 100 or fewer sequences. The final analysis
included 14 studies (Tables 1 and 2). There were seven studies with only
fecal samples (Ahn, Baxter, Brim, Hale, Wang, Weir, and Zeller), five
studies with only tissue samples (Burns, Dejea, Geng, Lu, Sanapareddy),
and two studies with both fecal and tissue samples (Chen and Flemer).
After curating the sequences, 1737 stool samples and 492 tissue samples
remained in the analysis {[}Tables 1 and 2{]}.

\textbf{\emph{Sequence Processing.}} Raw sequence data and metadata were
primarily obtained from the Sequence Read Archive (SRA) and dbGaP. Other
sequence and metadata were obtained directly from the authors (n = 4,
(17, 23, 25, 27)). Each dataset was processed separately using mothur
(v1.39.3) (45) using the default quality filtering methods for both 454
and Illumina sequence data. If it was not possible to use the defaults
because the sequences were trimmed too much, then the stated quality
cut-offs from the original study were used. Chimeric sequences were
identified and removed using VSEARCH (46). The curated sequences were
assigned to OTUs at 97\% similarity using the OptiClust algorithm (47)
and classified to the deepest taxonomic level that had 80\% support
using the naïve Bayesian classifier trained on the RDP taxonomy outline
(version 14, (48)).

\textbf{\emph{Community analysis.}} We calculated alpha diversity
metrics (i.e.~OTU richness, evenness, and Shannon diversity) for each
sample. Within each dataset, we ensured that the data followed a normal
distribution using power transformations. Using the transformed data, we
tested the hypothesis that individuals with normal colons, adenomas, and
carcinomas had significantly different alpha diversity metrics using
linear mixed-effect models. We also calculated the OR for each study and
metric by considering any value above the median alpha diversity value
to be positive. We measured the dissimilarity between individuals by
calculating the pairwise Bray-Curtis index and used PERMANOVA (49) to
test whether individuals with normal colons were significantly different
from those with adenomas or carcinomas. Finally, after binning sequences
into the deepest taxa that the naïve Bayesian classifier could calssify
the sequences, we quantified the ORs for individuals having an adenoma
or carcinoma and corrected for multiple comparisons using the
Benjamini-Hochberg method (50). Again, for each taxon, if the relative
abundance was greater than the median relative abundance for that taxon
in the study, the individual was considered to be positive.

\textbf{\emph{Random Forest classification analysis.}} To classify
individuals as having normal colons or tumors, we built Random Forest
classification models for each dataset and comparison using taxa with
significant ORs (after multiple comparison correction), all taxa, or
OTUs. Because no taxa were identified as having a significant OR
associated with adenomas using stool samples or tissue samples,
classification models based on OR data were not constructed to classify
individuals as having normal colons or adenomas. Within the training
dataset, 10-fold cross validation (5-fold cross validation for small
datasets) was used to build a model that was then evaluated on the
testing set. For the models constructed using the taxa with significant
ORs, the default mtry setting was used to train the model and this model
was tested on the other datasets in the meta-analysis. The reported AUC
values are the AUCs for the application of the model on the test sets.
For the OTU-based models, the dataset was split into training (80\% of
samples) and testing (20\%) sets and 10-fold cross validation (5-fold
cross validation for small datasets) on the training set was used to
generate the model for the testing set. The original 80/20 split and
fitting was repeated 100 times and the average AUC from these 100
repeats was reported. The Mean Decrease in Accuracy (MDA), a measure of
the importance of each taxon to the overall model was used to rank the
taxa used in each model. For all models, the default setting used was
\(\sqrt{p}\), where p is all the variables used in the respective model.
Normally, \(\sqrt{p}\) has been found to be what is chosen as the ideal
mtry after model tuneing (51).

\textbf{\emph{Statistical Analysis.}} All statistical analysis after
sequence processing utilized the R (v3.4.3) software package (52). For
OTU richness, evenness, and Shannon diversity analysis, values were
power transformed using the rcompanion (v1.11.1) package (53) and then
Z-score normalized using the car (v2.1.6) package (54). Testing for OTU
richness, evenness, and Shannon diversity differences utilized linear
mixed-effect models created using the lme4 (v1.1.15) package (55) to
correct for study, repeat sampling of individuals (tissue only), and 16S
hyper-variable region used. Odds ratios (OR) were analyzed using both
the epiR (v0.9.93) and metafor (v2.0.0) packages (56, 57) by assessing
how many individuals with and without disease were above and below the
overall median value within each specific study. OR significance testing
utilized the chi-squared test. Diversity differences measured by the
Bray-Curtis index utilized the creation of distance matrix and testing
with PERMANOVA executed with the vegan (v2.4.5) package (58). Random
Forest models were built using both the caret (v6.0.78) and randomForest
(v4.6.12) packages (59, 60). All figures were created using both ggplot2
(v2.2.1) and gridExtra (v2.3) packages (61, 62).

\textbf{\emph{Reproducible Methods.}} The code and analysis can be found
at
\url{https://github.com/SchlossLab/Sze_CRCMetaAnalysis_Microbiome_2017}.
Unless otherwise mentioned, the accession number of raw sequences from
the studies used in this analysis can be found directly in the
respective batch file in the GitHub repository or in the original
manuscript.

@@

\hypertarget{ref-keku_gastrointestinal_2015}{}
41. \textbf{Keku, T. O.}, \textbf{S. Dulal}, \textbf{A. Deveaux},
\textbf{B. Jovov}, and \textbf{X. Han}. 2015. The gastrointestinal
microbiota and colorectal cancer. American Journal of Physiology -
Gastrointestinal and Liver Physiology \textbf{308}:G351--G363.

\hypertarget{ref-vogtmann_epidemiologic_2016}{}
42. \textbf{Vogtmann, E.}, and \textbf{J. J. Goedert}. 2016.
Epidemiologic studies of the human microbiome and cancer. British
Journal of Cancer \textbf{114}:237--242.

\hypertarget{ref-kostic_genomic_2012}{}
43. \textbf{Kostic, A. D.}, \textbf{D. Gevers}, \textbf{C. S.
Pedamallu}, \textbf{M. Michaud}, \textbf{F. Duke}, \textbf{A. M. Earl},
\textbf{A. I. Ojesina}, \textbf{J. Jung}, \textbf{A. J. Bass},
\textbf{J. Tabernero}, \textbf{J. Baselga}, \textbf{C. Liu}, \textbf{R.
A. Shivdasani}, \textbf{S. Ogino}, \textbf{B. W. Birren}, \textbf{C.
Huttenhower}, \textbf{W. S. Garrett}, and \textbf{M. Meyerson}. 2012.
Genomic analysis identifies association of Fusobacterium with colorectal
carcinoma. Genome Research \textbf{22}:292--298.

>>>>

\textbf{\emph{Performance of models based on OTU relative abundances are
not significantly better than those based on taxa with significant
ORs.}} The previous models were based on relative abundance data where
sequences were classified to coarse taxonomic assignments
(i.e.~typically genus or family level). To determine whether model
performance improved with finer scale classification, we assigned
sequences to operational taxonomic units (OTUs) where the similarity
among sequences within an OTU was more than 97\%. We again found that
classification models built using all of the sequence data for a
community did a poor job of differentiating between subjects with normal
colons and those with adenomas (median AUC: 0.53 (0.37- 0.56)). However,
they did a good job of differentiating between subjects with normal
colons and those with carcinomas (median AUC: 0.71 (0.50- 0.90)). The
OTU-based models performed similarly to those constructed using the taxa
with significant ORs (one-tailed paired T-test; P-value=0.966) and those
using all taxa (one-tailed paired T-test; P-value=0.146) {[}Figure 4{]}.
Among the OTUs that had the highest mean decrease in accuracy for the
OTU-based models, we found that OTUs that affiliated with all of the 8
taxa that had a significant OR were within the top 10\% for at least one
study. This result was surprising as it indicated that a finer scale
classification of the sequences and thus a larger number of features to
select from, did not yield improved classification of the subjects.

\textbf{\emph{Generalizability of taxon-based models trained on one
dataset to the other datasets.}} Considering the good performance of the
Random Forest models trained using the relative abundance of taxa with
significant ORs and models trained using the relative abundance of all
taxa, we next asked how well the models would perform when given data
from a different cohort. For instance, if a model was trained using data
from the Ahn study, we wanted to know how well it would perform using
the data from the Baxter study. The models trained using the taxa with
significant ORs all had a higher median AUC than the models trained
using all of the taxa when tested on the other datasets {[}Figure 6 \&
S5{]}. As might be expected, the difference between the performance of
the modeling approaches appeared to vary with the size of the training
cohort (R\textsuperscript{2}=0.66) {[}Figure 6{]}. These data suggest
that given a sufficient number of subjects with normal colons and
carcinomas, Random Forest models trained using a small number of taxa
can accurately classify individuals from a different cohort.

@@

We performed a meta-analysis to identify and validate microbiota-base
biomarkers that could be used to classify individuals as having normal
colons or colonic tumors using fecal or tissue samples. To our surprise,
Random Forest classification models constructed to differentiate
individuals with normal colons from those with carcinomas using a subset
of the community performed well relative to models constructed using the
full communities. When we applied the models trained on each dataset to
the other datasets in our study, we found that the models trained using
the subset of the communities performed better than those using the full
communities. These models were trained using data in which sequences
were assigned to bacterial taxa using a classifier that typically
assigned sequences to the family or genus level. When we attempted to
improve the specificity of the classification by using an OTU-based
approach the resulting models performed as well as those constructed
using coarse taxonomic assignments. These results are significant
because they strengthen the growing literature indicating a role for the
colonic microbiota in tumorigenesis, as a potential tool as a
non-invasive diagnostic, and for assessing risk of disease and
recurrence (9, 12, 30).

Fine scale classification of sequences into OTUs did not improve our
classification models. This was also tested in earlier efforts to use
shotgun metagenomic data to classify individuals as having normal colons
or tumors; however, it was shown that analyses performed using shotgun
metagenomic data did not perform better than using 16S rRNA gene
sequencing data (31). We hypothesize that fine scale classification may
not result in better classification because distribution of microbiota
between individuals is patchy. In contrast, models using coarser
taxonomic assignments will pool the fine scale diversity, resulting in
less patchiness and better classification. Furthermore, the ability of
models trained using a subset of the community to outperform those using
the full community when testing the models on the other datasets may
also be a product of the patchiness of the human-associated microbiota.
The models based on the 8 taxa that had significant ORs used taxa that
were found in every study and tended to have higher relative abundances.
Similar to the OTU-based models, those models based on the full
community taxonomy assignments were still sensitive to the patchy
distribution of taxa. Regardless, it is encouraging that a collection of
8 taxa could reliably classify individuals as having carcinomas
considering the differences in cohorts, DNA extraction procedures,
regions of the 16S rRNA gene, and sequencing methods.

When used to classify individuals with carcinomas, the taxa with
significant ORs could not reliably classify individuals on their own
{[}Figure 3{]}. This result further supports the hypothesis that
carcinoma-associated microbiota have a patchy distribution. Two
individuals may have had the same classification, based on the relative
abundance of different populations within this group of 8 taxa. Although
these results only reflect associations with disease, it is tempting to
hypothesize that the patchiness is indicative of distinct mechanisms of
exacerbating tumorigenesis or that multiple taxa have the same mechanism
of exacerbating tumorigenesis. For example, strains of \emph{Escherichia
coli} and \emph{Fusobacterium nucleatum} have been shown to worsen
inflammation in mouse models of tumorigenesis (5, 6, 21). In contrast to
the patchiness of the taxa that were positively associated with
carcinomas, potentially beneficial taxa had a more consistent
association {[}Figure 6{]}. This result was particularly interesting
because members of these taxa (i.e. \emph{Ruminococcus} and
\emph{Clostridium XI} in fecal samples and \emph{Dorea} and
\emph{Blautia} in tissue) are thought to be beneficial due to their
involvement in production of anti-inflammatory short chain fatty acids
(32--34).

All of the adenoma classification models performed poorly, which is
consistent with previous studies (27, 30). However, the classification
results are at odds with results of the multitarget microbiota test
(MMT) from Baxter, et al. (12) who observed an AUC of 0.755 when the
test was applied to individuals with adenomas. There are two major
differences between the models generated in this meta-analysis and that
analysis. The MMT attempted to classify individuals as having a normal
colon or having colonic lesions (i.e.~adenomas or carcinomas) and not
adenomas alone. Further, the MMT incorporated fecal immunoglobulin test
(FIT) data while our models only used 16S rRNA gene sequencing data.
Because FIT data were not available for the other studies in our
meta-analysis, it was not possible to validate the MMT approach. The
ability to differentiate between individuals with and without adenomas
is an important problem since early detection of tumors is critical to
patient survivorship. However, it is possible that we might have been
able to detect differences in the bacterial community if individuals
with non-advanced and advanced adenomas were separated. This is a
clinically relevant distinction since advanced adenomas are at highest
risk of progressing to carcinomas. The initial changes of the microbiota
during tumorigenesis could be focal to where the initial adenoma
develops and would not be easily assessed using fecal samples from an
individual with non-advanced adenomas. Unfortunately, distinguishing
between individuals with advanced and non-advanced adenomas was not
possible in our meta-analysis since the studies did not provide the
clinical data needed to make that distinction.

Fecal samples represent a non-invasive approach to assess the structure
of the gut microbiota and are potentially useful for diagnosing
individuals as having colonic tumors. However, they do not reflect the
structure of the mucosal microbiota (35). Regardless, the taxa that were
the most important in the feces-based models overlapped with those from
the models trained using the data from unmatched and matched colon
tissue samples {[}Figure S3{]}. Mucosal biopsies are preferred for
focused mechanistic studies and have offered researchers the opportunity
to sample healthy and diseased tissue from the same individuals
(i.e.~matched) using each individual as their own control or in a
cross-sectional design (i.e.~unmatched). Because obtaining these samples
is invasive, carries risks to the individual, and is expensive, studies
investigating the structure of the mucosal microbiota generally have a
limited number of participants. Thus, it was not surprising that
tissue-based studies did not provide clearer associations between the
mucosal microbiota and the presence of tumors. Interestingly,
\emph{Fusobacterium}, which has received increased attention for its
potential role in tumorigenesis (6) was not consistently identified
across the studies in our meta-analysis. This could be due to the
relatively small number of individuals in the limited number of studies.
The classification models trained using the tissue-based data performed
well when tested with the training data (Figure S4), but performed
poorly when tested on the other tissue-associated datasets (Figure S5).
Disturbingly, taxa that are commonly associated with reagent
contamination (e.g. \emph{Novosphingobium}, \emph{Acidobacteria Gp2},
\emph{Sphingomonas}, etc.) were detected within the tissue datasets.
Such contamination is common in studies where there is relatively low
bacterial biomass (36). The lack of replication among the tissue-based
biomarkers may be a product of the relatively small number of studies
and individuals per study and possible reagent contamination.

Among the fecal sample data, we failed to identify several notable
populations that are commonly associated with carcinomas including an
enterotoxigenic strain of \emph{Bacteroides fragilis} (ETBF) and
\emph{Streptococcus gallolyticus} subsp. \emph{gallolyticus} (22, 24).
ETBF have been found in tumors in the proximal colon where they tend to
form biofilms (20, 37). Considering DNA from bacteria that are more
prevalent in the proximal colon may be degraded by the time it leaves
the body, it is not surprising that we failed to identify a significant
OR for \emph{Bacteroides} with carcinomas. In addition, since our
approach could only classify sequences to the genus level and there are
likely multiple \emph{Bacteroides} populations in the colon, it is
possible that sequences from ETBF and non-oncogenic \emph{Bacteroides}
were pooled. This would then reduce the OR between \emph{Bacterioides}
and whether an individual had carcinomas. It is also necessary to
distinguish between populations that are biomarkers for a disease and
those that are known to cause disease. Although the latter have been
shown to have a causative role, they may appear at low relative
abundance, be found in specific locations, or may have a highly patchy
distribution among affected individuals.

Meta-analyses are a useful tool in microbiome research because they can
demonstrate whether a result can be replicated and facilitate new
discoveries by pooling multiple independent investigations. There have
been several meta-analyses similar to this study that have sought
biomarkers for obesity (38--40), inflammatory bowel disease (39), and
colorectal cancer (28). Considering microbiome research is particularly
prone to hype and overgeneralization of results (41), these analyses are
critical. Meta-analyses are difficult to perform because the underlying
16S rRNA gene sequence data are not publicly available, metadata are
missing, incomplete, or vague, sequence data are of poor quality or
derived by non-standard approaches, and the original studies may be
significantly underpowered. Reluctance to publish negative results
(i.e.~the ``file drawer effect'') is also likely to skew our
understanding of the relationship between microbiota and disease. Better
attention to these specific issues will increase the reproducibility and
replicability of microbiota studies and make it easier to perform these
crucial meta-analyses. Moving forward, meta-analyses will be important
tools to help aggregate and find commonalities across studies when
investigating the microbiota in the context of a specific disease (28,
38--40).

@@

\textbf{\emph{Datasets.}} The studies used for this meta-analysis were
identified through the review articles written by Keku, et al. (42) and
Vogtmann, et al. (43). Additional studies, not mentioned in those
reviews were obtained based on the authors' knowledge of the literature.
Studies were included that used tissue or feces as their sample source
for 454 or Illumina 16S rRNA gene sequencing. A significant number of
studies (N=12) were excluded from the meta-analysis because they did not
have publicly available sequences, did not use 454 or Illumina
sequencing platforms, or did not have metadata that the authors were
able to share. We were able to obtain sequence data and metadata from
the following studies: Ahn, et al. (11), Baxter, et al. (12), Brim, et
al. (29), Burns, et al. (15), Chen, et al. (13), Dejea, et al. (20),
Flemer, et al. (17), Geng, et al. (19), Hale, et al. (27), Kostic, et
al. (44), Lu, et al. (26), Sanapareddy, et al. (25), Wang, et al. (14),
Weir, et al. (23), and Zeller, et al. (16). The Zackular (45) study was
excluded because the individuals studied were included within the larger
Baxter study (12). The Kostic study was excluded because after we
processed the sequences, all of the case samples had 100 or fewer
sequences. The final analysis included 14 studies (Tables 1 and 2).
There were seven studies with only fecal samples (Ahn, Baxter, Brim,
Hale, Wang, Weir, and Zeller), five studies with only tissue samples
(Burns, Dejea, Geng, Lu, Sanapareddy), and two studies with both fecal
and tissue samples (Chen and Flemer). After curating the sequences, 1737
fecal samples and 492 tissue samples remained in the analysis {[}Tables
1 and 2{]}.

\textbf{\emph{Sequence Processing.}} Raw sequence data and metadata were
primarily obtained from the Sequence Read Archive (SRA) and dbGaP. Other
sequence and metadata were obtained directly from the authors (n=4, (17,
23, 25, 27)). Each dataset was processed separately using mothur
(v1.39.3) using the default quality filtering methods for both 454 and
Illumina sequence data (46). If it was not possible to use the defaults
because the trimmed sequences were too short, then the stated quality
cut-offs from the original study were used. Chimeric sequences were
identified and removed using VSEARCH (47). The curated sequences were
assigned to OTUs at 97\% similarity using the OptiClust algorithm (48)
and classified to the deepest taxonomic level that had 80\% support
using the naïve Bayesian classifier trained on the RDP taxonomy outline
(version 14, (49)).

\textbf{\emph{Community analysis.}} We calculated alpha diversity
metrics (i.e.~OTU richness, evenness, and Shannon diversity) for each
sample. Within each dataset, we ensured that the data followed a normal
distribution using power transformations. Using the transformed data, we
tested the hypothesis that individuals with normal colons, adenomas, and
carcinomas had significantly different alpha diversity metrics using
linear mixed-effect models. We also calculated the OR for each study and
metric by considering any value above the median alpha diversity value
to be positive. We measured the dissimilarity between individuals by
calculating the pairwise Bray-Curtis index and used PERMANOVA (50) to
test whether individuals with normal colons were significantly different
from those with adenomas or carcinomas. Finally, after binning sequences
into the deepest taxa that the naïve Bayesian classifier could calssify
the sequences, we quantified the ORs for individuals having an adenoma
or carcinoma and corrected for multiple comparisons using the
Benjamini-Hochberg method (51). Again, for each taxon, if the relative
abundance was greater than the median relative abundance for that taxon
in the study, the individual was considered to be positive.

\textbf{\emph{Random Forest classification analysis.}} To classify
individuals as having normal colons or tumors, we built Random Forest
classification models for each dataset and comparison using taxa with
significant ORs (after multiple comparison correction), all taxa, or
OTUs. Because no taxa were identified as having a significant OR
associated with adenomas using stool samples or tissue samples,
classification models based on OR data were not constructed to classify
individuals as having normal colons or adenomas. Within the training
dataset, 10-fold cross validation (5-fold cross validation for small
datasets) was used to build a model that was then evaluated on the
testing set. For the models constructed using the taxa with significant
ORs, the default mtry setting was used to train the model and this model
was tested on the other datasets in the meta-analysis. The reported AUC
values are the AUCs for the application of the model on the test sets.
For the OTU-based models, the dataset was split into training (80\% of
samples) and testing (20\%) sets and 10-fold cross validation (5-fold
cross validation for small datasets) on the training set was used to
generate the model for the testing set. The original 80/20 split and
fitting was repeated 100 times and the average AUC from these 100
repeats was reported. The Mean Decrease in Accuracy (MDA), a measure of
the importance of each taxon to the overall model was used to rank the
taxa used in each model. For all models, the default setting used was
\(\sqrt{p}\), where p is all the variables used in the respective model.
Normally, \(\sqrt{p}\) has been found to be what is chosen as the ideal
mtry after model tuneing (52).

\textbf{\emph{Statistical Analysis.}} All statistical analysis after
sequence processing utilized the R (v3.4.3) software package (53). For
OTU richness, evenness, and Shannon diversity analysis, values were
power transformed using the rcompanion (v1.11.1) package (54) and
Z-score normalized using the car (v2.1.6) package (55). Testing for OTU
richness, evenness, and Shannon diversity differences utilized linear
mixed-effect models to correct for study, repeat sampling of individuals
(tissue only), and 16S rRNA gene sequence region used using the lme4
(v1.1.15) package (56). ORs were analyzed using both the epiR (v0.9.93)
and metafor (v2.0.0) packages (57, 58) by assessing how many individuals
with and without disease were above and below the overall median value
within each specific study. OR significance testing utilized the
chi-squared test. Community structure differences were calculated using
the Bray-Curtis dissimilarity index and PERMANOVA was used to test for
tumor-associated differences in structure with the vegan (v2.4.5)
package (59). Random Forest models were built using both the caret
(v6.0.78) and randomForest (v4.6.12) packages (60, 61). All figures were
created using both ggplot2 (v2.2.1) and gridExtra (v2.3) packages (62,
63).

\textbf{\emph{Reproducible Methods.}} The analysis code can be found at
\url{https://github.com/SchlossLab/Sze_CRCMetaAnalysis_mBio_2017}.
Unless otherwise mentioned, the accession number of raw sequences from
the studies used in this analysis can be found directly in the
respective batch file in the GitHub repository or in the original
manuscript.

@@

\hypertarget{ref-Hanage2014}{}
41. \textbf{Hanage, W. P.} 2014. Microbiology: Microbiome science needs
a healthy dose of scepticism. Nature. Springer Nature
\textbf{512}:247--248.

\hypertarget{ref-keku_gastrointestinal_2015}{}
42. \textbf{Keku, T. O.}, \textbf{S. Dulal}, \textbf{A. Deveaux},
\textbf{B. Jovov}, and \textbf{X. Han}. 2015. The gastrointestinal
microbiota and colorectal cancer. American Journal of Physiology -
Gastrointestinal and Liver Physiology \textbf{308}:G351--G363.

\hypertarget{ref-vogtmann_epidemiologic_2016}{}
43. \textbf{Vogtmann, E.}, and \textbf{J. J. Goedert}. 2016.
Epidemiologic studies of the human microbiome and cancer. British
Journal of Cancer \textbf{114}:237--242.

\hypertarget{ref-kostic_genomic_2012}{}
44. \textbf{Kostic, A. D.}, \textbf{D. Gevers}, \textbf{C. S.
Pedamallu}, \textbf{M. Michaud}, \textbf{F. Duke}, \textbf{A. M. Earl},
\textbf{A. I. Ojesina}, \textbf{J. Jung}, \textbf{A. J. Bass},
\textbf{J. Tabernero}, \textbf{J. Baselga}, \textbf{C. Liu}, \textbf{R.
A. Shivdasani}, \textbf{S. Ogino}, \textbf{B. W. Birren}, \textbf{C.
Huttenhower}, \textbf{W. S. Garrett}, and \textbf{M. Meyerson}. 2012.
Genomic analysis identifies association of Fusobacterium with colorectal
carcinoma. Genome Research \textbf{22}:292--298.
